---
title: "Forecast daily bike rental demand using time series models"
date: "`r Sys.Date()`"
output: html_document
author: "Jehangeer!"
---

# About Data Analysis Report

This RMarkdown file contains the report of the data analysis done for the project on forecasting daily bike rental demand using time series models in R. It contains analysis such as data exploration, summary statistics and building the time series models. The final report was completed on `r date()`. 

**Data Description:**

This dataset contains the daily count of rental bike transactions between years 2011 and 2012 in Capital bikeshare system with the corresponding weather and seasonal information.

**Data Source:** https://www.kaggle.com/datasets/marklvl/bike-sharing-dataset

**Relevant Paper:** 

Fanaee-T, Hadi, and Gama, Joao, 'Event labeling combining ensemble detectors and background knowledge', Progress in Artificial Intelligence (2013): pp. 1-15, Springer Berlin Heidelberg



# Load and explore the data

## Load data and install packages

```{r}
## Importing required packages

# install.packages("tidyverse")

library(tidyverse)

bikes_day <- read_csv("bike-sharing-dataset/day.csv")
bikes_hours <- read_csv("bike-sharing-dataset/hour.csv")

summary(bikes_day)
summary(bikes_hours)


```


## Describe and explore the data

```{r}

```



# Create interactive time series plots

```{r}
## Read about the timetk package
# ?timetk


```




# Smooth time series data

```{r}

```



# Decompse and access the stationarity of time series data

```{r}

```



# Fit and forecast time series data using ARIMA models

```{r}

```



# Findings and Conclusions































